---
title: "UVA Health Capstone"
author: "Rehan Merchant, Luke Ford, Avinaash Pavuloori, Ryan Williams"
date: "2/16/2022"
output: html_document
---


Thoughts:
1) We are going to need to do some preprocessing steps in the other r document such as changing data types but we should be able to do a majority of the modifications in this document.

Questions that we need to think about:
1) How are we going to deal with age because patients could have three different ages over the course of this three year period?
      I would recommend sorting by encounter number (if they are in chronological order) or by date if not and taking the age of the first encounter
2) How are we going to deal with deceased patients?
      Do we need to remove them from the data? Thoughts?
3) How do we deal with primary payer? (just their original one?)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(error=TRUE,        # Keep compiling upon error
                      collapse=FALSE,    # collapse by default
                      echo=TRUE,         # echo code by default
                      comment = "#>",    # change comment character
                      fig.width = 5.5,     # set figure width
                      fig.align = "center",# set figure position
                      out.width = "49%", # set width of displayed images
                      warning=TRUE,      # show R warnings
                      message=TRUE)      # show R messages
```

```{r}
# load required libraries
library(tidyverse)
library(ISLR)
library(boot)
library(caret)
library(MASS)
library(class)
library(lubridate)
```



# Original Data
```{r load-data, warning=FALSE, message=FALSE}
# loading patient data
original_data <- read.csv("full_data.csv") # change to match your directory
```

```{r}
# turn all empty cells into NA
original_data <- original_data %>% mutate_all(na_if, "")
```

## Basic Information
```{r}
# get size of the data frame
dim(original_data)
```

```{r}
# get names of columns
colnames(original_data)
```

```{r}
# preview of data
head(original_data)
```

# Data Wrangling
```{r}
data <- original_data[,-1] # drop the first row that is an old index
```

## Creating Subset of Necessary Rows
```{r}
# select the rows that are useful for the clustering and multinomial regression
data <- subset(data, select=c(ENC_NO, appt_dtm_mod, Patient_ID_mod, DEPARTMENT_NAME_EXTERNAL,
                              Encounter_Type_Status, CANCEL_REASON_NAME, 
                              Age_mod, Clrt_Sex, Clrt_Ethnic, PT_RACE, Clrt_Lang, Intrprtr_Need_YN,
                              Elixhauser, Enc_Primary_Payor, 
                              Visits_ED, Visits_EDtoIP, Visits_IP, PrimaryCare_Encounters, Visits_Completed, Visits_PCP, Visits_PCPDEPT,
                              ADULT_PALLIATIVE_CARE_REGISTRY_UVA, 
                              ANESTHESIA_RECORD_REGISTRY, 
                              ASTHMA_REGISTRY,BEWELL_REGISTRY_UVA, 
                              CANCER_POPULATION_REGISTRY,
                              CHRONIC_KIDNEY_DISEASE_REGISTRY,
                              CHRONIC_OPIOID_USE_REGISTRY_UVA,
                              CONGESTIVE_HEART_FAILURE_REGISTRY,
                              CONTROLLED_SUBSTANCE_REGISTRY_UVA,
                              COPD_REGISTRY,
                              CORONARY_ARTERY_DISEASE_REGISTRY,        
                              CYSTIC_FIBROSIS_REGISTRY,               
                              DIABETES_REGISTRY,                      
                              HYPERTENSION_REGISTRY,                  
                              ICU_STAY_REGISTRY,                      
                              INFLAMMATORY_BOWEL_DISEASE_UVA,         
                              INPATIENT_READMISSION_REGISTRY,          
                              JOINT_REPLACEMENT_REGISTRY_HIP_AND_KNEE,
                              LUNG_CANCER_SCREENING_REGISTRY_UVA,      
                              OBESITY_REGISTRY,                     
                              OSTEOPOROSIS_REGISTRY,                   
                              PEDIATRIC_OBESITY_REGISTRY,             
                              PREDIABETES_REGISTRY,                    
                              TOBACCO_REGISTRY,                      
                              WOUND_REGISTRY ))

# ENC_PROVIDER_TYPE, Is_Resdnt, PROC_NAME,CANCEL_INITIATOR, 
# Is_Walk_In_YN, Intrprtr_Need_YN, DEPARTMENT_NAME_EXTERNAL, Enc_Typ_Nme, Appt_Sts_Nme, 
# ADULT_PALLIATIVE_CARE_REGISTRY_UVA, ANESTHESIA_RECORD_REGISTRY, ASTHMA_REGISTRY, BEWELL_REGISTRY_UVA, CANCER_POPULATION_REGISTRY, CHRONIC_KIDNEY_DISEASE_REGISTRY,
# CHRONIC_OPIOID_USE_REGISTRY_UVA, CONGESTIVE_HEART_FAILURE_REGISTRY, CONTROLLED_SUBSTANCE_REGISTRY_UVA, COPD_REGISTRY, CORONARY_ARTERY_DISEASE_REGISTRY,
# CYSTIC_FIBROSIS_REGISTRY, DIABETES_REGISTRY, HYPERTENSION_REGISTRY, ICU_STAY_REGISTRY, INFLAMMATORY_BOWEL_DISEASE_UVA, INPATIENT_READMISSION_REGISTRY,
# JOINT_REPLACEMENT_REGISTRY_HIP_AND_KNEE, LUNG_CANCER_SCREENING_REGISTRY_UVA, OBESITY_REGISTRY, OSTEOPOROSIS_REGISTRY, PEDIATRIC_OBESITY_REGISTRY, PREDIABETES_REGISTRY,
# TOBACCO_REGISTRY, WOUND_REGISTRY,
#  PRINCIPLE_DIAGNOSIS, PRINCIPLE_DIAGNOSIS_DESC

dim(data)
```

```{r}
summary(data)
```

## Changing Data Types
```{r}
# ENC_NO - Encounter number shouldnt have to do anything
# sk_Fact_Pt_Enc_Clrt_mod - Not sure of meaning all are Na's so far
data$Patient_ID_mod <- as.factor(data$Patient_ID_mod)
# sk_Dim_Pt - Not sure of meaning all are Na's so far
#data$Contact_Date_mod <- ymd_hms(data$Contact_Date_mod) # guessing date appointment was made
#data$appt_dtm_mod <- ymd_hms(data$appt_dtm_mod) # guessing date and time appointment was made
#data$Adm_Dtm_mod <- ymd_hms(data$Adm_Dtm_mod) # guessing data of appointment
#data$Dsch_Dtm_mod <- ymd_hms(data$Dsch_Dtm_mod) # not sure what this is
#data$Is_Walk_In_YN <- as.factor(data$Is_Walk_In_YN) # walk in in need to figure what is yes
data$Intrprtr_Need_YN <- as.factor(data$Intrprtr_Need_YN) # need to figure out what factors mean
data$DEPARTMENT_NAME_EXTERNAL <- as.factor(data$DEPARTMENT_NAME_EXTERNAL) #department name
#data$Enc_Typ_Nme <- as.factor(data$Enc_Typ_Nme) # encounter type
#data$Appt_Sts_Nme <- as.factor(data$Appt_Sts_Nme) # appointment status
#data$adt_pt_cls <- as.factor(data$adt_pt_cls) # not sure yet
#data$ED_Disposition <- as.factor(data$ED_Disposition) # not sure yet
data$Encounter_Type_Status <- as.factor(data$Encounter_Type_Status) # what does the history factor mean
# PCP_Name - can probably get rid of this column
# data$PCP_PROVIDER_TYPE <-  as.factor(data$PCP_PROVIDER_TYPE)
# ENC_PROVIDER_NAME
#data$ENC_PROVIDER_TYPE <- as.factor(data$ENC_PROVIDER_TYPE)
#data$Is_Resdnt <- as.factor(data$Is_Resdnt)
#data$PROC_NAME <- as.factor(data$PROC_NAME) #check for miss spellings
# APPT_LENGTH
#data$ROOMED_DTTM_mod <- ymd_hms(data$ROOMED_DTTM_mod) # not sure what this is maybe time they entered room
#data$SIGNIN_DTTM_mod <- ymd_hms(data$SIGNIN_DTTM_mod) # im guessing time signed in
# TIME_TO_ROOM_MINUTES
# TIME_IN_ROOM_MINUTES
# CYCLE_TIME_MINUTES
data$CANCEL_REASON_NAME <- as.factor(data$CANCEL_REASON_NAME) # maybe look to subtract non-patient reasons from total cancels
#data$CANCEL_INITIATOR <- as.factor(data$CANCEL_INITIATOR) # look above
data$Clrt_Ethnic <- as.factor(data$Clrt_Ethnic)
data$Clrt_Lang <- as.factor(data$Clrt_Lang)
# Age_mod need to modify age. Im guessing 90+ is reprsented as 900, but check with angela
#data$Death_Dte_mod <- ymd_hms(data$Death_Dte_mod)
data$Clrt_Sex <- as.factor(data$Clrt_Sex)
data$PT_RACE <- as.factor(data$PT_RACE)
#data$last_pcp_prov_visit_dtm_mod <- ymd_hms(data$last_pcp_prov_visit_dtm_mod)
#data$last_pcp_dept_visit_dtm_mod <- ymd_hms(data$last_pcp_dept_visit_dtm_mod)
# months_from_pcp_prov_visit # 999 appears to be NA
# months_from_pcp_deptv_visit_dtm # 999 appears to be NA
data$ADULT_PALLIATIVE_CARE_REGISTRY_UVA <- as.factor(data$ADULT_PALLIATIVE_CARE_REGISTRY_UVA) # what do these mean
data$ANESTHESIA_RECORD_REGISTRY <- as.factor(data$ANESTHESIA_RECORD_REGISTRY)
data$ASTHMA_REGISTRY <- as.factor(data$ASTHMA_REGISTRY)
data$BEWELL_REGISTRY_UVA <- as.factor(data$BEWELL_REGISTRY_UVA)
data$CANCER_POPULATION_REGISTRY <- as.factor(data$CANCER_POPULATION_REGISTRY)
data$CHRONIC_KIDNEY_DISEASE_REGISTRY <- as.factor(data$CHRONIC_KIDNEY_DISEASE_REGISTRY)
data$CHRONIC_OPIOID_USE_REGISTRY_UVA <- as.factor(data$CHRONIC_OPIOID_USE_REGISTRY_UVA)
data$CONGESTIVE_HEART_FAILURE_REGISTRY <- as.factor(data$CONGESTIVE_HEART_FAILURE_REGISTRY)
data$CONTROLLED_SUBSTANCE_REGISTRY_UVA <- as.factor(data$CONTROLLED_SUBSTANCE_REGISTRY_UVA)
data$COPD_REGISTRY <- as.factor(data$COPD_REGISTRY)
data$CORONARY_ARTERY_DISEASE_REGISTRY <- as.factor(data$CORONARY_ARTERY_DISEASE_REGISTRY)
data$CYSTIC_FIBROSIS_REGISTRY <- as.factor(data$CYSTIC_FIBROSIS_REGISTRY)
data$DIABETES_REGISTRY <- as.factor(data$DIABETES_REGISTRY)
data$HYPERTENSION_REGISTRY <- as.factor(data$HYPERTENSION_REGISTRY)
data$ICU_STAY_REGISTRY <- as.factor(data$ICU_STAY_REGISTRY)
data$INFLAMMATORY_BOWEL_DISEASE_UVA <- as.factor(data$INFLAMMATORY_BOWEL_DISEASE_UVA)
data$INPATIENT_READMISSION_REGISTRY <- as.factor(data$INPATIENT_READMISSION_REGISTRY)
data$JOINT_REPLACEMENT_REGISTRY_HIP_AND_KNEE <- as.factor(data$JOINT_REPLACEMENT_REGISTRY_HIP_AND_KNEE)
data$LUNG_CANCER_SCREENING_REGISTRY_UVA <- as.factor(data$LUNG_CANCER_SCREENING_REGISTRY_UVA)
data$OBESITY_REGISTRY <- as.factor(data$OBESITY_REGISTRY)
data$OSTEOPOROSIS_REGISTRY <- as.factor(data$OSTEOPOROSIS_REGISTRY)
data$PEDIATRIC_OBESITY_REGISTRY <- as.factor(data$PEDIATRIC_OBESITY_REGISTRY)
data$PREDIABETES_REGISTRY <- as.factor(data$PREDIABETES_REGISTRY)
data$TOBACCO_REGISTRY <- as.factor(data$TOBACCO_REGISTRY)
data$WOUND_REGISTRY <- as.factor(data$WOUND_REGISTRY)
# Elixhauser
data$Enc_Primary_Payor <- as.factor(data$Enc_Primary_Payor)
# Visits_ED
# Visits_EDtoIP
# Visits_IP
# Visits_ED_2019
# Visits_EDtoIP_2019
# Visits_IP_2019
# Visits_ED_2020
# Visits_EDtoIP_2020
# Visits_IP_2020
# Visits_ED_2021
# Visits_EDtoIP_2021
# Visits_IP_2021
# PrimaryCare_Encounters
# Visits_Completed
# Visits_PCP
# Visits_PCPDEPT
#data$PRINCIPLE_DIAGNOSIS <- as.factor(data$PRINCIPLE_DIAGNOSIS)
# PRINCIPLE_DIAGNOSIS_DESC
```


## Modifications To Entire Data Set
```{r}
# modifications
data <- data %>% dplyr::arrange(appt_dtm_mod) %>% # sorting by appointment date so when grouping by we get initial age, primary payor etc
  mutate(Age_mod = ifelse(Age_mod == 900,90, Age_mod), # taking 900 value which represnts any one over 90 and changing it to 90, so all 90+ year olds will be represented by the number 90 [Confirmed]
         Elixhauser = ifelse(is.na(Elixhauser), 0, Elixhauser)) # people with perfect elixhauser scores were being represented by NAs, mutatuing the NAs to equal 0 instead. [need to confrim]

# modification for primary_insurance

data$insurance_buckets <- ifelse(grepl("MEDICARE", data$Enc_Primary_Payor),"Medicare",ifelse(grepl("MEDICAID", data$Enc_Primary_Payor),"Medicaid",ifelse(grepl("Self Pay", data$Enc_Primary_Payor),"Self Pay","Private")))
data$insurance_buckets <- as.factor(data$insurance_buckets)

# modification for language / maybe interpreter needed - need to figure out a better way to link if a interpreter is needed might need to create another table and join if we want it
data$lang_buckets <- ifelse(grepl("English",data$Clrt_Lang),"English",ifelse(grepl("Spanish",data$Clrt_Lang),"Spanish","Other"))
data$lang_buckets <- as.factor(data$lang_buckets)

# modification for cancel reason name, buckets were based on smaller data set, will have to adjust with the bigger data set with new features
# NA, Other, Patient, Provider
data$cancel_reason_bucket <- ifelse(grepl("Change in Medical Necessity",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Deceased",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Error",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Facility Cancelled",data$CANCEL_REASON_NAME),"Provider",ifelse(grepl("Home Health - Rescheduled with another provider",data$CANCEL_REASON_NAME),"Patient",ifelse(grepl("Home Health Provider",data$CANCEL_REASON_NAME),"Provider",ifelse(grepl("Hospitalized",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Information Added/Revised",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Lacking Required Insurance Referral or Authorization",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Patient Left Without Being Seen",data$CANCEL_REASON_NAME),"Patient",ifelse(grepl("Patient's Request",data$CANCEL_REASON_NAME),"Patient",ifelse(grepl("Patient's Request - Cancelled Via Automated Reminder System",data$CANCEL_REASON_NAME),"Patient",ifelse(grepl("Patient's Request - Transportation Issues",data$CANCEL_REASON_NAME),"Patient",ifelse(grepl("Provider",data$CANCEL_REASON_NAME),"Provider",ifelse(grepl("Provider Unavailable",data$CANCEL_REASON_NAME),"Provider",ifelse(grepl("Referring Physician Cancelled",data$CANCEL_REASON_NAME),"Provider",ifelse(grepl("Short notice cancel/reschedule by patient",data$CANCEL_REASON_NAME),"Patient",ifelse(grepl("Weather or Facility Issue",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("COVID",data$CANCEL_REASON_NAME),"COVID",ifelse(grepl("Coordinating Appointment",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Patient too ill",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Rescheduled to Telemedicine Visit",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Cancel",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Scheduled from Wait List",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Patient's request - Seeing Non-UVA Provider/Wait List Is Too Long",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Understaffed",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Patient Technical Issues",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Patient",data$CANCEL_REASON_NAME),"Patient",ifelse(grepl("Cancelled via MyChart",data$CANCEL_REASON_NAME),"Patient",ifelse(grepl("Home Health - unable to reach patient",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("HIS Only - Duplicate MRN",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("Appointment Scheduled in Error",data$CANCEL_REASON_NAME),"Other",ifelse(grepl("At Capacity",data$CANCEL_REASON_NAME),"Other",NA)))))))))))))))))))))))))))))))))


data$cancel_reason_bucket <- as.factor(data$cancel_reason_bucket)

#modification of race

data$race_bucket <- ifelse(grepl("Asian",data$PT_RACE),"Asian",ifelse(grepl("Not Available",data$PT_RACE),"Unknown",ifelse(grepl("Hispanic",data$PT_RACE),"Other",ifelse(grepl("Indian",data$PT_RACE),"Asian",ifelse(grepl("Unknown",data$PT_RACE),"Unknown",ifelse(grepl("Other",data$PT_RACE),"Other",ifelse(grepl("UNKNOWN",data$PT_RACE),"Unknown",ifelse(grepl("Black",data$PT_RACE),"Black",ifelse(grepl("White",data$PT_RACE),"White","Unknown")))))))))

# confirm these are the right representations
```

```{r}
#change registry NA to 0

data[, 10:50][is.na(data[, 10:50])] <- 0
```

```{r}
dim(data)
```

```{r}
#to delete
cancel_table <- data %>% count(CANCEL_REASON_NAME) %>% arrange(n)
cancel_table

race_table <- data %>% count(PT_RACE) %>% arrange(n)
race_table
```

```{r}
data[0:500,] %>% arrange(desc(Encounter_Type_Status))
```

```{r}
#delete?
data_test  <- data[-which(data$appt_dtm_mod == ""), ]
data_test %>% dplyr::filter(Patient_ID_mod == 466155)  %>%  arrange(desc(Enc_Primary_Payor), appt_dtm_mod)
```

# Data Set with Factors Important for Multinomial Regression
Create seperate data set with features for the multinomial regression - We can add in the registries later if we think they could be important
```{r}
# Need to look at this further. For some reason the first data_mult section isnt working

#data_mult <- data[-which(data$appt_dtm_mod == ""), ] # | data$Enc_Primary_Payor == "" ), ] # create a second set of data # removes all empty dates and primary payor
#data2 <- data2[!is.na(data2$Elixhauser), ] # removes NAs for Elixhauser (produces 27,424 vs 34,957 distinct patient IDs so might not want to remove them) # dont think we want to remove them
data_mult <- distinct(data, Patient_ID_mod, .keep_all = TRUE)
data_mult <- data_mult %>% dplyr::select(Patient_ID_mod, Elixhauser, insurance_buckets, Age_mod, Clrt_Sex, PT_RACE, Clrt_Ethnic, lang_buckets, cancel_reason_bucket, race_bucket, ADULT_PALLIATIVE_CARE_REGISTRY_UVA, 
                              ANESTHESIA_RECORD_REGISTRY, 
                              ASTHMA_REGISTRY,BEWELL_REGISTRY_UVA, 
                              CANCER_POPULATION_REGISTRY,
                              CHRONIC_KIDNEY_DISEASE_REGISTRY,
                              CHRONIC_OPIOID_USE_REGISTRY_UVA,
                              CONGESTIVE_HEART_FAILURE_REGISTRY,
                              CONTROLLED_SUBSTANCE_REGISTRY_UVA,
                              COPD_REGISTRY,
                              CORONARY_ARTERY_DISEASE_REGISTRY,        
                              CYSTIC_FIBROSIS_REGISTRY,               
                              DIABETES_REGISTRY,                      
                              HYPERTENSION_REGISTRY,                  
                              ICU_STAY_REGISTRY,                      
                              INFLAMMATORY_BOWEL_DISEASE_UVA,         
                              INPATIENT_READMISSION_REGISTRY,          
                              JOINT_REPLACEMENT_REGISTRY_HIP_AND_KNEE,
                              LUNG_CANCER_SCREENING_REGISTRY_UVA,      
                              OBESITY_REGISTRY,                     
                              OSTEOPOROSIS_REGISTRY,                   
                              PEDIATRIC_OBESITY_REGISTRY,             
                              PREDIABETES_REGISTRY,                    
                              TOBACCO_REGISTRY,                      
                              WOUND_REGISTRY) #might want to add in the specific registries later on and if interpreter is needed
head(data_mult)
dim(data_mult)
```

Possible Features to include for multinomial regression
• elixhauser (do we want initial or average)
• primary insurance (think we will only be able to take first, any other thoughts) (initial might be better for predicting new patients but current might be better for clustering moving forward)


• age
• gender
• ethnicity
• race
• interpreter needed
• registries

```{r}
patient_id <- data$Patient_ID_mod
unique_patient_id <- unique(patient_id)
length(unique_patient_id)
```

```{r}
#to delete?
data %>% dplyr::filter(Patient_ID_mod == 791225) %>% arrange(desc(Elixhauser))
```

```{r}
summary(data_mult)
```

# Data Set with Factors for Clustering

## Modifications for Clustering Data Set

```{r}

```


### Reshaping Data
```{r}
#data to reshape - takes roughly 3.5 minutes
#CANCEL_REASON_NAME
#data_reshape <- data.frame(model.matrix(~cancel_reason_bucket-1, data=data))
#data_reshape <- subset(data_reshape, select = -c(cancel_reason_bucket))
#data_clust <- cbind(data,data_reshape)

data_clust <- data

#data_reshape <- data.frame(model.matrix(~CANCEL_REASON_NAME-1, data=data))
#data_reshape <- subset(data_reshape, select = -c(CANCEL_REASON_NAME))
#data_clust <- cbind(data,data_reshape)

#Encounter_Type_Status
#data_reshape2 <- data.frame(model.matrix(~Encounter_Type_Status-1, data=data))
#data_reshape2 <- subset(data_reshape2, select = -c(Encounter_Type_Status))
#data_clust <- cbind(data_clust,data_reshape2)
#data_clust
```


```{r}
colnames(data_clust)
```

```{r}
data_clust <- subset(data_clust, select=c(ENC_NO, appt_dtm_mod, Patient_ID_mod, Encounter_Type_Status,
                              Visits_ED, Visits_EDtoIP, Visits_IP, PrimaryCare_Encounters, Visits_Completed, Visits_PCP, Visits_PCPDEPT))
                     
                     
#                              CANCEL_REASON_NAMECancelled.via.MyChart, CANCEL_REASON_NAMEHome.Health...unable.to.reach.patient, CANCEL_REASON_NAMELacking.Required.Insurance.Referral.or.Authorization, CANCEL_REASON_NAMEPatient.Left.Without.Being.Seen,
#                              CANCEL_REASON_NAMEPatient.s.Request...Cancelled.Via.Automated.Reminder.System, CANCEL_REASON_NAMEPatient.s.Request...Transportation.Issues, CANCEL_REASON_NAMEShort.notice.cancel.reschedule.by.patient,
#                              CANCEL_REASON_NAMEPatient, CANCEL_REASON_NAMEPatient.Technical.Issues))
```


### Feature Engineering
```{r}
# create the group_by
patient <- data_clust %>% group_by(Patient_ID_mod) # each patient is distinct

#select(Patient_ID_mod, PrimaryCare_Encounters, Visits_PCP, Visits_ED, Visits_EDtoIP, Visits_IP, CANCEL_REASON_NAMEPatient.s.Request
# create the new grouped data set
data_clust <- patient  %>% dplyr::select(ENC_NO, appt_dtm_mod, Patient_ID_mod, Encounter_Type_Status,
                              Visits_ED, Visits_EDtoIP, Visits_IP, PrimaryCare_Encounters, Visits_Completed, Visits_PCP, Visits_PCPDEPT) %>% 
                              #,CANCEL_REASON_NAMECancelled.via.MyChart, CANCEL_REASON_NAMEHome.Health...unable.to.reach.patient, CANCEL_REASON_NAMELacking.Required.Insurance.Referral.or.Authorization, CANCEL_REASON_NAMEPatient.Left.Without.Being.Seen,
                              #CANCEL_REASON_NAMEPatient.s.Request...Cancelled.Via.Automated.Reminder.System, CANCEL_REASON_NAMEPatient.s.Request...Transportation.Issues, CANCEL_REASON_NAMEShort.notice.cancel.reschedule.by.patient,
                              #CANCEL_REASON_NAMEPatient, CANCEL_REASON_NAMEPatient.Technical.Issues) %>% 
  
  dplyr::summarise(PCPDEPT = mean(Visits_PCPDEPT), ED = mean(Visits_ED), IP = mean(Visits_IP), EDtoIP = mean(Visits_EDtoIP), PC_Encounters = mean(PrimaryCare_Encounters)) %>% 
                  # , Visits_Completed = mean(Visits_Completed), CANCEL_REASON_NAMECancelled.via.MyChart = sum(CANCEL_REASON_NAMECancelled.via.MyChart), CANCEL_REASON_NAMEHome.Health...unable.to.reach.patient = sum(CANCEL_REASON_NAMEHome.Health...unable.to.reach.patient),CANCEL_REASON_NAMELacking.Required.Insurance.Referral.or.Authorization = sum(CANCEL_REASON_NAMELacking.Required.Insurance.Referral.or.Authorization),CANCEL_REASON_NAMEPatient.Left.Without.Being.Seen = sum(CANCEL_REASON_NAMEPatient.Left.Without.Being.Seen), CANCEL_REASON_NAMEPatient.s.Request...Cancelled.Via.Automated.Reminder.System = sum(CANCEL_REASON_NAMEPatient.s.Request...Cancelled.Via.Automated.Reminder.System),CANCEL_REASON_NAMEPatient.s.Request...Transportation.Issues = sum(CANCEL_REASON_NAMEPatient.s.Request...Transportation.Issues), CANCEL_REASON_NAMEShort.notice.cancel.reschedule.by.patient = sum(CANCEL_REASON_NAMEShort.notice.cancel.reschedule.by.patient), CANCEL_REASON_NAMEPatient = sum(CANCEL_REASON_NAMEPatient), CANCEL_REASON_NAMEPatient.Technical.Issues = sum(CANCEL_REASON_NAMEPatient.Technical.Issues)) %>% 
  
  dplyr::mutate(PC_Visits = PCPDEPT, ED_Visits = (ED + EDtoIP)) %>% 
                
                #, Non_Visit_Encounters = PC_Encounters - PCPDEPT, Visit_Completed_Percentage = Visits_Completed / (Visits_Completed + CANCEL_REASON_NAMECancelled.via.MyChart + CANCEL_REASON_NAMEHome.Health...unable.to.reach.patient + CANCEL_REASON_NAMELacking.Required.Insurance.Referral.or.Authorization + CANCEL_REASON_NAMEPatient.Left.Without.Being.Seen + CANCEL_REASON_NAMEPatient.s.Request...Cancelled.Via.Automated.Reminder.System + CANCEL_REASON_NAMEPatient.s.Request...Transportation.Issues +CANCEL_REASON_NAMEShort.notice.cancel.reschedule.by.patient +CANCEL_REASON_NAMEPatient +  CANCEL_REASON_NAMEPatient.Technical.Issues)) %>% 
  
  dplyr::select(Patient_ID_mod, PC_Visits, PC_Encounters, ED_Visits) #, Non_Visit_Encounters) 

#data_clust$ED_Percentage[is.na()] <- 0

# replace index with patient ID
#clustering_data <- remove_rownames(clustering_data)
#clustering_data <- column_to_rownames(clustering_data, 'Patient_ID_mod')

#data_clust$ED_Percentage[is.na(data_clust$ED_Percentage)] <- 0 # changes NA values in ED_Percentage to 0s. The NA is caused  when there are 0 Total visits.

head(data_clust)                                                                                                                                            
```

Feature Engineering 
• Total_Visits = Visits_PCPDEPT + Visits_ED + Visits_IP + Visits_EDtoIP # Dr. Babit only cares about Visits_PCPDEPT
• ED_Percentage = (Visits_ED + Visits_EDtoIP) / Total_Visits
• Non_Visit_Encounters = PrimaryCare_Encounters - Visits_PCPDEPT

• Visit_Completed_Percentage = Visits_Completed / (Visits_Completed + CANCEL_REASON_NAMEShort.notice.cancel.reschedule.by.patient + CANCEL_REASON_NAMEPatient.s.Request + CANCEL_REASON_NAMEPatient.s.Request...Transportation.Issues + CANCEL_REASON_NAMEPatient.s.Request...Cancelled.Via.Automated.Reminder.System + ANCEL_REASON_NAMEPatient.Left.Without.Being.Seen)


• CANCEL_REASON_NAME (used to determine if cancel reason was on patients fault)
• Encounter_Type_Status (are certain encounters less resource intensive then others)
• ENC_PROVIDER_TYPE (does it matter resource wise who they are seeing)

```{r}
#to delete
#data_clust %>% dplyr::arrange(ED_Percentage)
```

```{r}
#to delete?
data %>% dplyr::select(Patient_ID_mod,  Visits_PCPDEPT, Visits_ED, Visits_IP, Visits_EDtoIP) %>%  dplyr::filter(Patient_ID_mod ==  153)
```


```{r}
#to delete
data_clust$ED_Percentage[is.na(data_clust$ED_Percentage)] <- 0
data_clust %>%  dplyr::filter(is.na(ED_Percentage))
```


```{r}
head(patient)
```


```{r}
#to delete?
data %>% dplyr::filter(Patient_ID_mod == 4470807) %>% dplyr::arrange(Encounter_Type_Status)
```


```{r}
#to delete?
patient %>% dplyr::filter(Patient_ID_mod == 4470807) %>% dplyr::arrange(Encounter_Type_Status)
```


# Merge Clustering Data and Multinomial Data

```{r}
data_final <- merge(data_mult, data_clust, by = "Patient_ID_mod")
head(data_final)
```

```{r}
write.csv(data_final, "cleaned_data.csv",row.names = FALSE)
```


To delete below?

## Modeling


```{r}
scaled_data <- scale(clustering_data) # scale the data
```



k-medioud
https://www.statology.org/k-medoids-in-r/
```{r}
library(factoextra)
library(cluster)

#scaled_data
```

determine optimal number of clusters
```{r}
#fviz_nbclust(scaled_data, pam, method = 'wss', k.max = 8)
```

```{r}
#gap_stat <- clusGap(scaled_data, FUN = pam, K.max = 20, B = 50)

#fviz_gap_stat(gap_stat)


```



```{r}
pamx <- pam(data_clust, 3, metric = 'euclidean', stand = TRUE) # Try True and remove scaling before
```

```{r}
summary(pamx)
```

```{r}
plot(pamx)
```



# K-Mediod with 4 Clusters looking at Primary Care Visits and Emergency Room Visits
```{r}
pamx$medoids
```


```{r}
fviz_cluster(pamx, scaled_data, geom = 'point', repel = FALSE, show.clust.cent = TRUE)
```


## Next Steps
Place the cluster back into the data and perform multinomial regression to determine what predictors are the best predictors for each grouping (age, gender, insurance provider, etc)